---
title: "Summary of TerraBio eDNA Results"
author: "Karen Dyson"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of eDNA data

This should have the eDNA reads

Total number of species variants


## Key species questions and indicators

### Key species present in our sample

### Biodiversity Indicator 4 (ABF-KPI-6): Number of keystone/priority species due to intervention.

We found _____ species present in cocoa fields (intervention) that were not present in pasture (counterfactual). These species include:

Additionally, we found _____ species present in pasture not present in cocoa fields. These species include:

______ species were present in both cocoa and pasture fields.



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
